---
title: "Factors of Alcoholism"
subtitle: "in teens"
author: "Team Violet"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(knitr)
library(xaringanthemer)
library(palmerpenguins)
library(showtext)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
```

```{r}
style_xaringan(
  title_slide_background_image = "presentation/img/confetti.jpg"
)
```


class: center, middle

## A statement of the overall goal / research question

The goal of our research will be to identify the academic and social factors that lead to alcoholism in teenagers. We will then construct a model to identify the likelihood of teenagers to suffer from alcoholism.

class: inverse, center, middle

# Section title

---

# Hello World

```{r load_data}
student_por <- read.csv("../data/student-por.csv")

ncol(student_por)

student_por <- student_por %>% mutate(Talc=(2/7*(Walc)+5/7*(Dalc)))

```

```{r EDA_Aamir}

student_por %>% group_by(sex) %>% summarise(mean=mean(Talc))

student_por %>% 
  ggplot(aes(x=Talc, y=health)) +
  geom_jitter() + geom_smooth()

student_por %>% 
  ggplot(aes(x=Medu, y=Talc)) + 
  geom_jitter() + 
  geom_smooth()

student_por %>% 
  ggplot(aes(x=Fedu, y=Talc)) + 
  geom_jitter() + 
  geom_smooth()
  
student_por %>%
  ggplot(aes(x=Talc, fill=sex)) +
  geom_histogram(binwidth = .25)

student_por %>%
  ggplot(aes(x=age, y=Talc)) +
  geom_jitter() +
  geom_smooth()

student_por %>%
  group_by(Mjob) %>%
  summarise(mean=mean(Talc))

student_por %>%
  group_by(Fjob) %>%
  summarise(mean=mean(Talc))

```
```{r linear_reg}
lin_model <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Talc ~ school +
        sex +
        age + 
        address + 
        famsize + 
        Pstatus + 
        Medu +
        Fedu +
        Mjob + 
        Fjob + 
        reason +
        guardian +
        traveltime +
        studytime +
        failures +
        schoolsup +
        famsup +
        paid +
        activities +
        nursery +
        higher +
        internet +
        romantic +
        famrel +
        freetime +
        goout +
        health +
        absences, data = student_por)

line_tidy <- tidy(lin_model) %>% 
  filter(p.value < 0.05)

line_tidy
```
```{r new_model}
lin_model2 <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Talc ~ school +
        age + 
        failures + 
        freetime + 
        sex +
        Fjob + 
        Mjob +
        reason + 
        studytime +
        famsup +
        goout +
        health +
        absences, data = student_por)
```




```{r resid}
aug <- augment(lin_model2$fit)

aug %>%
  ggplot(mapping = aes(x = .fitted, y = .resid)) +
  geom_jitter() + 
  geom_hline(yintercept = 0)
```
Note limitation - prediction confidence may be too low, std error artificially low


Visualizing school

```{r school, eval = TRUE}
student_por_school <- student_por %>% 
  select(school, Talc) %>% 
  group_by(school) %>%
  summarise(average_talc = mean(Talc))


student_por_school %>%
  ggplot(mapping = aes(x = school, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By School", 
    x = "School", 
    y = "Average Total Alcohol Consumption"
  )

```

There does not appear to be a siginificant difference between schools.
---
Visualizing Sex

```{r sex, eval = TRUE}
student_por_sex <- student_por %>% 
  select(sex, Talc) %>% 
  group_by(sex) %>%
  summarise(average_talc = mean(Talc))


student_por_sex %>%
  ggplot(mapping = aes(x = sex, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Sex", 
    x = "Sex", 
    y = "Average Total Alcohol Consumption"
  )
```

Males seem to drink more, on average, than females. 
---
Visualizing age

```{r age, eval = TRUE}
student_por_age <- student_por %>% 
  select(age, Talc)

  
student_por_age %>%
  ggplot(mapping = aes(x = age, y = Talc)) +
  geom_jitter() +
  labs(
    title = "Age vs Total Alcohol Consumption", 
    x = "Age", 
    y = "Total Alcohol Consumption"
  )
```

There does not seem to be a correlation between age and total alcohol consumption. 

---
Visualizing Address

```{r address, eval = TRUE}
student_por_address <- student_por %>% 
  select(address, Talc) %>% 
  group_by(address) %>%
  summarise(average_talc = mean(Talc))


student_por_address %>%
  ggplot(mapping = aes(x = address, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Address", 
    x = "Address", 
    y = "Average Total Alcohol Consumption"
  )
```
There does not seem to be a significant difference between average alchol consumption between rural and urban highschools. 

---
Visualizing Family Size

```{r famsize, eval = TRUE}
student_por_famsize <- student_por %>% 
  select(famsize, Talc) %>% 
  group_by(famsize) %>%
  summarise(average_talc = mean(Talc))


student_por_famsize %>%
  ggplot(mapping = aes(x = famsize, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Family Size", 
    x = "Family Size", 
    y = "Average Total Alcohol Consumption"
  )
```
There does not appear to be a significant difference between average alcohol consumption and family size. 

---
Visualizing Parent's Cohabitation Status

```{r Pstatus, eval = TRUE}
student_por_Pstatus <- student_por %>% 
  select(Pstatus, Talc) %>% 
  group_by(Pstatus) %>%
  summarise(average_talc = mean(Talc))


student_por_Pstatus %>%
  ggplot(mapping = aes(x = Pstatus, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Parental Cohabitation Status", 
    x = "Cohabitation Status", 
    y = "Average Total Alcohol Consumption"
  )
```
Highschoolers with parents who live together have a higher average total alcohol consumption value, than those whose parents live apart. 
---
Visualizing Mother's Education 

```{r Medu, eval = TRUE}
student_por_Medu <- student_por %>% 
  select(Medu, Talc)


student_por_Medu %>%
  ggplot(mapping = aes(x = Medu, y = Talc)) +
  geom_jitter() +
  labs(
    title = "Mothers Education vs Average Total Alcohol Consumption", 
    x = "Mother's Education", 
    y = "Total Alcohol Consumption"
  )
```
There does not seem to be a relationship between mother's education and total alcohol consumption.
---
Visualizing Father's Education 

```{r Fedu, eval = TRUE}
student_por_Fedu <- student_por %>% 
  select(Fedu, Talc)


student_por_Fedu %>%
  ggplot(mapping = aes(x = Fedu, y = Talc)) +
  geom_jitter() +
  labs(
    title = "Father's Education vs Average Total Alcohol Consumption", 
    x = "Father's Education", 
    y = "Total Alcohol Consumption"
  )
```
There does not seem to be a relationship between father's education and total alcohol consumption.

---
Visualizing Mother's Job

```{r Mjob, eval = TRUE}
student_por_Mjob <- student_por %>% 
  select(Mjob, Talc) %>% 
  group_by(Mjob) %>%
  summarise(average_talc = mean(Talc))


student_por_Mjob %>%
  ggplot(mapping = aes(x = Mjob, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Mother's Job", 
    x = "Mother's Job", 
    y = "Average Total Alcohol Consumption"
  )
```
High schoolers with mothers who work in health care have the lowest average alcohol consumption, while those with mothers who work in civil services have the highest. 

---
Visualizing Father's Job

```{r Fjob, eval = TRUE}
student_por_Fjob <- student_por %>% 
  select(Fjob, Talc) %>% 
  group_by(Fjob) %>%
  summarise(average_talc = mean(Talc))


student_por_Fjob %>%
  ggplot(mapping = aes(x = Fjob, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Father's Job", 
    x = "Father's Job", 
    y = "Average Total Alcohol Consumption"
  )
```
High schoolers whose fathers are teachers have the lowest total alcohol consumption while those with fathers in civil service jobs have the highest. 

---
Visualizing Reason Job

```{r reason, eval = TRUE}
student_por_reason <- student_por %>% 
  select(reason, Talc) %>% 
  group_by(reason) %>%
  summarise(average_talc = mean(Talc))


student_por_reason %>%
  ggplot(mapping = aes(x = reason, y = average_talc)) +
  geom_col() +
  labs(
    title = "Average Total Alcohol Consumption By Reason for Joining School", 
    x = "Reason", 
    y = "Average Total Alcohol Consumption"
  )
```
High schoolers who choose their school based on reputation have the lowest total alohol consumption while those who choose for reasons besides courses, closeness to home, and reputation, have the highest. 
---
- Click the `Knit` button to compile your presentation

- Make sure to commit and push all resulting files to your GitHub repo

---

class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
